\propertitle

### 2019-12-04 maxNt_v6.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-12-04T17:01:22+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal population\\ from subpopulation data {\large\color{myred}[draft]}%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

\mynote{Rewrite!}

@@

The second assumption, typical of maximum-entropy in general, concerns the
form of the probability distribution for the long-run frequencies ensuing
from the first assumption. It is taken to be either an \enquote{entropic
  prior} \citep{skillingetal1984,rodriguez1991,neumann2007}, which takes
into account the multiplicity of long-run frequencies through a coefficient
$L$, and a reference distribution, or a prior for a model with sufficient
statistics, which expresses the sufficiency of the moments under study
\citep{portamana2017}. This assumption is crucial for obtaining the
maximum-entropy method with the Shannon entropy. A different density leads
to alternative maximum-entropy methods, for example based on the Burg
entropy \citep{burg1975} if a Dirichlet density is used
\citep{jaynes1986d_r1996}.

@@

The space of possible frequency distributions for the larger population has
$N$ dimensions, however, and probability distributions in high dimensions
have counter-intuitive properties. For example, the mode or mean can have
\emph{atypical} features when compared with the majority of other probable
points of the probability space. The question then is \emph{which features
  of the maximum-entropy frequency distribution are typical of the majority
  of plausible frequency distributions?} Preliminary studies using a full
probability calculation (to be published separately) show that several
semi-quantitative features of the maximum-entropy distribution are indeed
typical. For the first data set, for example, most frequency distributions
have two high-frequency regimes of activity, as in the upper panel of
\fig~\ref{fig:plots_5mom}, with roughly the same heights and almost the
same locations, within a few percentages of normalized total activity. Each
of high-frequency regimes, though, may consist of two or three very close
modes rather than one.

@@

\mynote{add this above}We could obviously consider pre-data beliefs different
from~\eqref{eq:entropic_prior}, for example one quantified by a Dirichlet
distribution (which is equivalent to the above but with $\yF$ and $\yR$
switched), or a uniform distribution in $\yF$-space. Would these lead to
markedly different post-data beliefs? A full-fledged probabilistic analysis
shows that the three typical features listed above still appear with these
different initial beliefs. They are therefore robust.

@@

\mynote{cite  \cite[p.~440]{maxwell1873c}}

We have presented a procedure to construct the most plausible frequency
distribution of population-averaged activities of a population of neurons,
given the recording about a small sample thereof. This procedure combines
the maximum-entropy method and basic identities from sampling theory. From
the application to two real data sets we saw that the frequency
distributions obtained with our procedure can have features very different
from the one measured in the sample, such as multiple modes. This procedure
can also be used with moment constraints of different order -- means,
population-averaged pairwise correlations, or higher-order correlations --
thus giving an approximate assessment of the informational sufficiency of
specific subsets of moments. In fact, we saw that the application of
maximum-entropy only at the sample level leads to misleading results about
this kind of sufficiency questions.

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-12-04T20:38:52+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal population\\ from subpopulation data {\large\color{myred}[draft]}%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

@@

The second assumption, typical of maximum-entropy in general, concerns the
form of the probability distribution for the long-run frequencies ensuing
from the first assumption. It is taken to be either an \enquote{entropic
  prior} \citep{skillingetal1984,rodriguez1991,neumann2007}, which takes
into account the multiplicity of long-run frequencies, or a prior for a
model with sufficient statistics, which expresses the sufficiency of the
moments under study \citep{portamana2017}. This assumption is crucial for
obtaining the maximum-entropy method with the Shannon entropy. A different
density leads to alternative maximum-entropy methods, for example based on
the Burg entropy \citep{burg1975} if a Dirichlet density is used
\citep{jaynes1986d_r1996}. The entropic prior depends on a reference
distribution and on a coefficient $L$ which determines the sharpness of the
prior's peak. The computations of the preceding sections were insensitive
to the reference distribution in their significant digits. We used three
biologically reasonable alternatives: a flat distribution, one linearly
decreasing with $\yA$, and a normal one with a broad peak at low
activities.

@@

The space of possible frequency distributions for the larger population has
$N$ dimensions, however, and probability distributions in high dimensions
have counter-intuitive properties. For example, the mode or mean can have
\emph{atypical} features when compared with the majority of other probable
points of the probability space. The question then is \emph{which features
  of the maximum-entropy frequency distribution are typical of the majority
  of plausible frequency distributions?} Preliminary studies using a full
probability calculation (to be published separately) show that several
semi-quantitative features of the maximum-entropy distribution are indeed
typical. For the first data set, for example, most frequency distributions
have two high-frequency regimes of activity, as in the upper panel of
\fig~\ref{fig:plots_5mom}, with roughly the same heights and almost the
same locations, within a few percentages of normalized total activity. Each
of high-frequency regimes, though, may consist of two or three very close
modes rather than one. The same preliminary studies also indicate that
these typical features remain with prior choices in the second assumption
above, for example using a Dirichlet prior instead of an entropic one.

@@

@@

The reason why we record neurons from a brain region is to have an idea of
the activity of the other neurons in that region, similarly to survey
sampling. If we thought that \emph{precisely} the recorded ones were
exceptions or completely unrelated to the others surrounding the recording
probe, our recording would serve little purpose. Probabilistic methods that
allow us to make direct inferences about the larger population are
therefore useful. But they become complex and sensitive if we ask for very
detailed inferences.

We have presented a method that allows us to make direct inferences about a
larger population from very small samples. Its inferences regard only the
frequencies of the total activity of the larger population, but such
limited scope makes the method computationally very cheap.

Yet, and most important, the method leads to inferences that are not at all
trivial and not easily discernible by looking at the sample. We showed this
by applying the method to two real data sets.

By \enquote{inference} we mean the quantification of a degree of belief
about possible observations, based on specific assumptions; not the
response of an oracle. The method gives us a \emph{forecast} based on those
assumptions, and therefore has two main uses, just like any other
probabilistic method. If our assumptions are only hypotheses under testing,
then the subsequent confirmation or confutation of the forecast will
increase or decrease our confidence in those hypotheses. If we strongly
trust our assumptions, on the other hand, then we rely on the forecast and
even use it in place of actual observations, especially if the latter are
difficult to make. (In fact, we typically shift from the first to the
second use.) In the preceding sections we have given examples of hypotheses
that could be explored with the method. The method itself rests on
assumptions, discussed in \sect~\ref{sec:assumptions}, that can either be
under test or relied upon, depending on the case.

Some approximations implicit in the method make its numerical results
semi-quantitative with respect to an analysis based on the full probability
calculus. The results should not be off by more than a few percentages for
the normalized total activity and for the frequency density. This
imprecision, however, is compensated by the extreme computational cheapness
of the method, far less costly than a full probabilistic analysis. The
method is therefore very convenient at least for essaying hypotheses. We
hope to present an exact quantitative comparison with the full
probabilistic approach in a future publication.

% \mynote{cite  \cite[p.~440]{maxwell1873c}}

% We have presented a procedure to construct the most plausible frequency
% distribution of population-averaged activities of a population of neurons,
% given the recording about a small sample thereof. This procedure combines
% the maximum-entropy method and basic identities from sampling theory. From
% the application to two real data sets we saw that the frequency
% distributions obtained with our procedure can have features very different
% from the one measured in the sample, such as multiple modes. This procedure
% can also be used with moment constraints of different order -- means,
% population-averaged pairwise correlations, or higher-order correlations --
% thus giving an approximate assessment of the informational sufficiency of
% specific subsets of moments. In fact, we saw that the application of
% maximum-entropy only at the sample level leads to misleading results about
% this kind of sufficiency questions.
